---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# spatialLIBD

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Travis build status](https://travis-ci.org/LieberInstitute/spatialLIBD.svg?branch=master)](https://travis-ci.org/LieberInstitute/spatialLIBD)
<!-- badges: end -->

The `spatialLIBD` package contains functions for:

* Accessing the spatial transcriptomics data from the LIBD Human Pilot project ([code on GitHub](https://github.com/LieberInstitute/HumanPilot)) generated with the Visium platform from 10x Genomics. The data is retrieved from [Bioconductor](http://bioconductor.org/)'s `ExperimentHub`.
* Visualizing the spot-level spatial gene expression data and clusters.
* Inspecting the data interactively either on your computer or through [jhubiostatistics.shinyapps.io/spatialLIBD/](https://jhubiostatistics.shinyapps.io/spatialLIBD/).


## Installation instructions

Get the latest stable `R` release from [CRAN](http://cran.r-project.org/). Then install `spatialLIBD` using the following code:

```{r 'install', eval = FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("spatialLIBD")
```


## Citation

Below is the citation output from using `citation('spatialLIBD')` in R. Please 
run this yourself to check for any updates on how to cite __spatialLIBD__.

```{r 'citation', eval = requireNamespace('spatialLIBD')}
citation('spatialLIBD')
```


## Code of conduct


Please note that the `spatialLIBD` project is released with a
  [Contributor Code of Conduct](CODE_OF_CONDUCT.md).
  By contributing to this project, you agree to abide by its terms.
  
